---
title: "Reproducible Research Course Project"
author: "Sam"
date: "4/26/2022"
output:
  pdf_document: default
  html_document: default
---

## Reproducible Research Course Project

## Reading the dataset and processing the data

```{r}
activity <- read.csv("activity.csv")
str(activity)
activity$date <- as.Date(activity$date)
head(activity)
```

## Dropping NA rows 

```{r}
library(tidyverse)
activity <- drop_na(activity, steps)
sum(is.na(activity$steps))
```

## Histogram of the total number of steps taken each day

```{r}
step_total <- aggregate(steps~date,activity,sum)
head(step_total)
dim(step_total)

hist(step_total$steps)
```

## Mean and median number of steps taken each day

```{r}
mean_steps <- aggregate(steps~date,activity,mean)
mean_steps

median_steps <- aggregate(steps~date,activity,median)
median_steps
```

## Time series plot of the average number of steps taken based on 5-minute interval

```{r}
library(ggplot2)
tsdata <- aggregate(steps~interval,activity,mean)
head(tsdata)
dim(tsdata)
ggplot(data = tsdata, aes(x = interval, y = steps)) + geom_line()
```

**The time interval between 750 - 1000 had the most number of steps.**

## Code to describe and show a strategy for imputing missing data

```{r}
imputed_activity <- activity
```

## Replacing all the NA values with the mean of steps w.r.t their corresponding intervals

```{r}
imputed_activity[which(is.na(imputed_activity$steps))] <- tsdata$steps[imputed_activity$interval == tsdata$interval]

sum(is.na(imputed_activity))

head(imputed_activity)
dim(imputed_activity)

imputed_activity$steps <- ifelse(is.na(imputed_activity$steps), tsdata$steps, imputed_activity$steps)
sum(is.na(imputed_activity$steps))
```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r}
hist(imputed_activity$steps)
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
library(lubridate)
imputed_activity$day <- weekdays(activity$date)
```

## Creating dataframes grouped by weekends and weekdays to create a panel plot

```{r}
weekend_data <- subset(imputed_activity, day == c("Saturday", "Sunday"), select = c(steps, date, interval, day))
head(weekend_data)
unique(weekend_data$day)
```

```{r}
weekday_data <- subset(imputed_activity, day == c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), select = c(steps, date, interval, day))
head(weekday_data)
unique(weekday_data$day)
```

## Panel plot

```{r}
par(mfrow=c(2,1))
plot(y = weekend_data$steps, x = weekend_data$interval, type = "l")
plot(y = weekday_data$steps, x = weekday_data$interval, type = "l")
```
